Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-04-17 Moltres.tex
moves periods after citations

Molten salt reactors are a reactor concept that received considerable interest
in the 1950s and 60s with development of the Aircraft Reactor Experiment and
later the \gls{MSRE} at \gls{ORNL}. With the inclusion of the \gls{MSR} in the
Generation-IV reactor designs, the reactor concept has gained renewed research
interest in the past decade. Some of the advantages offered by the \gls{MSR}
include lower fissile inventories, no radiation damage constraint on attainable
fuel burn-up, no spent nuclear fuel, no requirement to fabricate and handle
solid fuel, and a homogeneous isotopic composition of fuel in the
reactor. \cite{noauthor_gen_2008} These characteristics give the \gls{MSR} a
potentially unique capability for actinide burning and extension of fuel
resources.

\gls{MSR} modeling efforts are focused on description of steady-state and
transient behavior. Krepel et. al. extended the in-house light-water reactor
diffusion code DYN3D to consider drift of delayed neutron precursors alongside
the reactor temperature profile, re-casting the extended code as
DYN3D-MSR. \cite{krepel_dyn3d-msr_2007} DYN3D-MSR was compared against
experimental \gls{MSRE} data and was then used to simulate local fuel channel
blockages as well as local temperature perturbations. In a similar vein, Kophazi
et. al. used iterative coupling between in-house three dimensional neutronic and
one-dimensional heat conduction models DALTON and THERM to analyze the normal
operational state of the \gls{MSRE} and some channel-blocking-incident
transients. \cite{kophazi_development_2009} Additions in the Kophazi model
included entrance effects of heat transfer coefficients as well as thermal
coupling between fuel channels through moderator heat conduction. More recently,
Cammi et. al. performed a 2D-axisymmetric single-channel analysis of the
\gls{MSBR} using the commercial finite element package COMSOL
Multiphysics. \cite{cammi_multi-physics_2011} Advances in this work included
solution of the fuel salt velocity field, use of heterogeneous group constants
in fuel and moderator regions, and employment of a software package (COMSOL)
intrinsically designed for coupled multi-physics simulation.

@@

The model is implemented using the \gls{MOOSE} \cite{gaston_physics-based_2015} application code
Moltres. \cite{lindsay_arfc/moltres_nodate} Building on the massively
parallelizable \gls{MOOSE} framework allows Moltres to be run on super-computing
platforms like \gls{NCSA}'s Blue Waters. For some three-dimensional simulations
(results not discussed here), the number of elements in the mesh and total
number of degrees of freedom exceed one million and ten million respectively. To
handle problems of this size, Moltres has been run on up to 608 cores. However,
reducing the problem dimension from three to two and using a structured mesh
that can be much more coarse in the axial direction allows problem solution to
be carried out on a single core in under five minutes.

@@

Reactor composition is based on the enriched uranium design for the \gls{MSRE}
and is given in \cref{table:comp}. \cite{robertson_msre_1965} Group constants
are generated with either Serpent \cite{leppanen_serpent_2015} or Scale
\cite{dehart_reactor_2011}. Temperature vs group constant interpolation tables
are constructed separately for fuel and moderator regions.

>>>>

Molten salt reactors are a reactor concept that received considerable interest
in the 1950s and 60s with development of the Aircraft Reactor Experiment and
later the \gls{MSRE} at \gls{ORNL}. With the inclusion of the \gls{MSR} in the
Generation-IV reactor designs, the reactor concept has gained renewed research
interest in the past decade. Some of the advantages offered by the \gls{MSR}
include lower fissile inventories, no radiation damage constraint on attainable
fuel burn-up, no spent nuclear fuel, no requirement to fabricate and handle
solid fuel, and a homogeneous isotopic composition of fuel in the
reactor \cite{noauthor_gen_2008}. These characteristics give the \gls{MSR} a
potentially unique capability for actinide burning and extension of fuel
resources.

\gls{MSR} modeling efforts are focused on description of steady-state and
transient behavior. Krepel et. al. extended the in-house light-water reactor
diffusion code DYN3D to consider drift of delayed neutron precursors alongside
the reactor temperature profile, re-casting the extended code as
DYN3D-MSR \cite{krepel_dyn3d-msr_2007}. DYN3D-MSR was compared against
experimental \gls{MSRE} data and was then used to simulate local fuel channel
blockages as well as local temperature perturbations. In a similar vein, Kophazi
et. al. used iterative coupling between in-house three dimensional neutronic and
one-dimensional heat conduction models DALTON and THERM to analyze the normal
operational state of the \gls{MSRE} and some channel-blocking-incident
transients \cite{kophazi_development_2009}. Additions in the Kophazi model
included entrance effects of heat transfer coefficients as well as thermal
coupling between fuel channels through moderator heat conduction. More recently,
Cammi et. al. performed a 2D-axisymmetric single-channel analysis of the
\gls{MSBR} using the commercial finite element package COMSOL
Multiphysics \cite{cammi_multi-physics_2011}. Advances in this work included
solution of the fuel salt velocity field, use of heterogeneous group constants
in fuel and moderator regions, and employment of a software package (COMSOL)
intrinsically designed for coupled multi-physics simulation.

@@

The model is implemented using the \gls{MOOSE} \cite{gaston_physics-based_2015} application code
Moltres \cite{lindsay_arfc/moltres_nodate}. Building on the massively
parallelizable \gls{MOOSE} framework allows Moltres to be run on super-computing
platforms like \gls{NCSA}'s Blue Waters. For some three-dimensional simulations
(results not discussed here), the number of elements in the mesh and total
number of degrees of freedom exceed one million and ten million respectively. To
handle problems of this size, Moltres has been run on up to 608 cores. However,
reducing the problem dimension from three to two and using a structured mesh
that can be much more coarse in the axial direction allows problem solution to
be carried out on a single core in under five minutes.

@@

Reactor composition is based on the enriched uranium design for the \gls{MSRE}
and is given in \cref{table:comp} \cite{robertson_msre_1965}. Group constants
are generated with either Serpent \cite{leppanen_serpent_2015} or Scale
\cite{dehart_reactor_2011}. Temperature vs group constant interpolation tables
are constructed separately for fuel and moderator regions.
